---
title: "exercises"
output: html_document
date: "2025-11-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1 Write a R program to take input from the user (name and age) and display the values. Also print the version of R installation
```{r}
name <- readline(prompt = "Enter your name: ")
```

#Take input for age and convert to numeric
```{r}
age_conv <- readline(prompt = "Enter your age: ")
age <- as.numeric(age_conv)
```

#Display the collected name and age
```{r}
print(paste("Hi,", name, "! You are", age, "years old."))
```

#Display the version of R installation
```{r}
getRversion()
```

#2 Write a R program to create a sequence of numbers from 20 to 50 and find the mean of numbers from 20 to 60 and sum of numbers from 51 to 91.
```{r}
print(seq(20,50))
print(mean(20:60))
print(sum(51:91))
```

#3 Write a R program to create a vector which contains 10 random integer values between -50 and +50.
```{r}
v <- sample(x = -50:50, size = 10)
print(v)
```

#4 Write a R program to find the maximum and the minimum value of a given vector.
```{r}
v = c(55:90)
print(v)
print(paste("Maximum value of vector v:", max(v)))
print(paste("Minimum value of vector v:", min(v)))
```

#5 Write a R program to create three vectors a,b,c with 3 integers. Combine the three vectors to become a 3×3 matrix where each column represents a vector. Print the content of the matrix.  
```{r}
x = c(36,67,82)
y = c(44,74,51)
z = c(94,34,68)
  
m <- cbind(x, y, z)
  
print("Content of matrix m:")
print(m)
```

#6 Write a R program to create a 5 x 4 matrix , 3 x 3 matrix with labels and fill the matrix by rows and 2 × 2 matrix with labels and fill the matrix by columns.
```{r}
m1 = matrix(31:50, nrow=5, ncol=4)
print("5 x 4 matrix:")

value1 = c(23,14,25,75,45,73,97,67,63)
row_n = c("1","2","3")
col_n = c("A","B","C")
m2 = matrix(value1, nrow=3, ncol=3, byrow=TRUE, dimnames=list(row_n, col_n))
print("3 x 3 matrix:")
print(m2)
  
value2 = c(5:8)
row_n2 = c("1","2")
col_n2 = c("Y","Z")
m3 = matrix(value2, nrow=2, ncol=2, byrow=FALSE, dimnames=list(row_n2, col_n2))
print("2 x 2 matrix:")
print(m3)
```

#7 Write a R program to create an array, passing in a vector of values and a vector of dimensions. Also provide names for each dimension.
```{r}
a = array(
  34:64,
  dim = c(4, 4, 2),
  dimnames = list(
  c("1","2","3","4"),
  c("A","B","C","D"),
  c("part1","part2")
                  )
          )
print(a)
```

#8 Write a R program to create a list of elements using vectors, matrices and a functions. Print the content of the list.
```{r}
li = list(
  c(3, 5, 8, 7, 9, 12),
  letters, #predefined vector
  matrix(c(4, 6, 8, 10), nrow = 2),
  tan
  )
  
print("Content of the list:")
print(li)
```

#9 Write a R program to create a simple bar plot of five subjects marks.
```{r}
marks = c(9, 15, 4, 17)
  
# Create a bar plot with the specified marks
  barplot(marks,
          
# Title of the bar plot
          main = "Comparing marks of 5 subjects",
          
# Label the x-axis as "Marks"
          xlab = "Marks",
          
# Label the y-axis as "Subject"
          ylab = "Subject",
          
# Set the names of the bars
          names.arg = c("Oceanog.", "Program.", "Legislat.", "Genetics"),
          
# Set the color of the bars to dark red
          col = "darkblue",
          
# Set the bars to vertical
          horiz = FALSE)
```
  
#10 Write a R program to create a Dataframes which contain details of 5 employees and display the details.
```{r}
employees = data.frame(
  name = c("Alice Mars", "Sarah Pyten", "Harry Worls", "Ron Height", "Kate Sheli"),
  gender = c("F", "F", "M", "M", "F"),
  age = c(24,34,42,58,47),
  role = c("Supervisor","Manager","Salesman","CEO","Assistant"),
  phone = c("524564","987489","763876","527565","958365")
  )

print ("Details of the employees:")
print (employees)
```

#11 Write a R program to compute sum, mean and product of a given vector 
```{r}
v = c(24,64,74,80)

print (paste ("The sum result is:", sum(v)))
print (paste ("The mean result is:", mean(v)))
print (paste ("The product result is:", prod(v)))
```

