---
title: "exercises"
output: html_document
date: "2025-11-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1 Write a R program to create an array using four given columns, three given rows, and two given tables and display the content of the array.
```{r}
array1 = array(10:40, dim=c(3,4,2))

print(array1)
```

#2 Write a R program to create a two-dimensional 5x3 array of sequence of even integers greater than 50.
```{r}
x <- array(seq(from = 50, length.out = 15, by = 2), c(5, 3))

print("Content of the array, sequence of even integers greater than 50:")
print(x)
```

#3 Write a R program to find row and column index of maximum and minimum value in a given matrix
```{r}
m1 = matrix(c(1:16), nrow = 4, byrow = TRUE)

print("Original Matrix:")
print(m1)

result = which(m1 == max(m1), arr.ind=TRUE)
print("Row and column of maximum value of matrix m1:")
print(result)

result = which(m1 == min(m1), arr.ind=TRUE)
print("Row and column of minimum value of matrix m1:")
print(result)
```

#4 Write a R program to concatenate two given matrices of same column but different rows.
```{r}
a = matrix(25:36, ncol=3, , byrow = TRUE)
b = matrix(38:58, ncol=3, , byrow = TRUE)

print("Matrix-1")
print(a)

print("Matrix-2")
print(b)

result = rbind(a, b)

print("Matrices concatenated:")
print(result)
```

#5 Write a R program to get the length of the first two vectors of a given list.
```{r}
listA = list(a1 = 11:20, a2 = "R Programming", a3 = "PHP")

print("Original list:")
print(listA)

print("Length of the vector 'a1' and 'a2' of listA:")
print(length(listA$a1))
print(length(listA$a2))
```

#6 Write a R program to find all elements of a given list that are not in another given list.
```{r}
l1 = list("a", "b", "c")
l2 = list("A", "B", "C", "a", "b", "c")

print("Original lists:")
print(l1)
print(l2)

print("All elements of l2 that are not in l1:")
setdiff(l2, l1)
```

#7 Write a R program to create a list containing a vector, a matrix and a list and remove the second element.
```{r}
list_data <- list(c("Pink","Blue","White"), matrix(c(4,8,3,2,7,5), nrow = 2),
                  list("C##", "HTML", "Ruby"))
print("List:")
print(list_data)

print("Remove the second element of the list:")
list_data[2] = NULL

print("New list:")
print(list_data)
```

#8 Write a R program to save the information of a data frame in a file and display the information of the file. 
# Create a data frame named 'exam_data' with columns: name, score, attempts, and qualify
```{r}
exam_data = data.frame(
  name = c('Marie', 'Dylan', 'Kathy', 'John', 'Emma', 'Michele', 'Matt', 'Laury', 'Keven', 'Josh'),
  score = c(12.5, 9, 16.5, 12, 9, 20, 14.5, 13.5, 8, 19),
  attempts = c(1, 3, 2, 3, 2, 3, 1, 1, 2, 1),
  qualify = c('yes', 'no', 'yes', 'no', 'no', 'yes', 'yes', 'no', 'no', 'yes')
)

print("Original dataframe:")
print(exam_data)

save(exam_data, file = "data.rda")

load("data.rda")

file.info("data.rda")
```

#9 Write a R program to create a data frame using two given vectors and display the duplicated elements and unique rows of the said data frame. 
```{r}
x = c(10, 20, 10, 10, 40, 50, 20, 30)
y = c(10, 30, 10, 20, 0, 50, 30, 30)

print("Original data frame:")

xy = data.frame(x, y)
print(xy)

print("Duplicate elements of the said data frame:")
print(xy[duplicated(xy), ])

print("Unique rows of the said data frame:")
print(unique(xy))
```

#10 Write a R program to find elements which are present in two given data frames.
```{r}
x = c("a", "b", "c", "d", "e")
y = c("d", "e", "f", "g")

print("Original Dataframes")

print(x)
print(y)

print("Elements which are present in both dataframes:")
result = intersect(x, y)
print(result)
```

#11 Write a R program to create an ordered factor from data consisting of the names of months. 
```{r}
month_v = c("March","April","January","November","January","September","October","September","November","August","February","January","November","November","February","May","August","February","July","December","August","August","September","November","September","February","April")

print("Original vector:")
print(month_v)

f = factor(month_v)

print("Ordered factors of vector month_v:")
print(f)

print(table(f))
```

#12 Write a R program to add 3 to each element in a given vector. Print the original and new vector.
```{r}
vec = c(1, 2, NULL, 3, 4, NULL)

print("Original vector:")
print(vec)

new_vec = (vec + 3)[(!is.na(vec)) & vec > 0]

print("New vector:")
print(new_vec)
```

#13 Write a R program to create a vector and find the length and the dimension of the vector.
```{r}
vec = c(1, 3, 5, 7, 9)

print("Original vector:")
print(vec)

print("Dimension of the vector:")
print(dim(vec)) #returns NULL for a 1D vector

print("Length of the vector:")
print(length(vec))
```

